---
title: "LASSO&COX"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,include=FALSE}
library(rgl)
library(TCGAbiolinks)
library(TCGA2STAT)
library(ggplot2)
library(survival)
library(ggpubr)
library(survminer)
library(survivalROC)
library(plotROC)
library(Matrix)
library(glmnet)
library(polspline)
library(rms)
library(dplyr)
library(VIM)
library(reshape2)
library(MASS)
library(Heatplus)
library(reshape2)
library(RColorBrewer)
library(ConsensusClusterPlus)
library(sigclust)
library(pheatmap)
library(doMC)
library(tsne)
library(caret)
library(magrittr)
library(tidyverse)
library(plyr)
library(gplots)
library(DESeq2)
library(factoextra)
library(timeROC)
```

### Glioma Prognosis Analysis using LASSO and COX

```{r}
data_dir <- "../Data/survival_data"
genes_path <- file.path(data_dir,"Glioma_40_genes.txt")
gt44 <- read_tsv(genes_path,col_names=FALSE,show_col_types=FALSE)$X1

### This part splits the file Glioma_survival.csv into 7:3 training dataset and testing dataset
Glioma_subtypes_path <- file.path(data_dir,"Glioma_survival.csv")
Giioma_subs <- read.csv(Glioma_subtypes_path, header = TRUE)
Glioma_subs <- na.omit(Glioma_subs)

sub <- sample(1:nrow(Glioma_subs),round(nrow(Glioma_subs)*0.7))
training <- Glioma_subs[sub,]
testing <- Glioma_subs[-sub,]

# write.csv(training,file = "Glioma_training.csv",quote=FALSE,row.names = FALSE)
# write.csv(testing,file = "Glioma_testing.csv",quote=FALSE,row.names = FALSE)
# training_path <- file.path(data_dir,"Glioma_training.csv")
# testing_path <- file.path(data_dir,"Glioma_testing.csv")
# read training and testing dataset that are saved
# training <- read.csv(training_path,header = TRUE)
# testing <- read.csv(testing_path,header = TRUE)

# extract 44 genes
training <- training[,colnames(training) %in% c(gt44,"patient","OS","Status","pam50")]
testing <- testing[,colnames(testing) %in% c(gt44,"patient","OS","Status","pam50")]

# Preprocess the training data
pp = preProcess(training[,!colnames(training) %in% c("patient","Status","OS","Original.Subtype")], method=c("scale"))
dt = predict(pp, training[,!colnames(training) %in% c("patient","Status","OS","Original.Subtype")])
x <- data.matrix(dt)
# change survival time into year unit
dt$OS = training$OS/12
dt$Status = training$Status

# Process the testing data
pp = preProcess(testing[,!colnames(testing) %in% c("patient","Status","OS","Original.Subtype")], method=c("scale"))
pptesting = predict(pp, testing[,!colnames(testing) %in% c("patient","Status","OS","Original.Subtype")])
# change survival time into year unit
pptesting$OS = testing$OS/12
pptesting$Status = testing$Status
```

### Using LASSO to get the marker gene

```{r}
# Set up dependent variable y, package survival time and survival state (survival data)
y <- data.matrix(Surv(as.double(dt$OS),as.double(dt$Status)))
# sum(y<0)
fit <-glmnet(x,y,family = "cox",alpha = 1)
# plot(fit,xvar="lambda",label=T)

#cross validation
set.seed(1234)
fitcv <- cv.glmnet(x,y,family="cox", alpha=1,nfolds=10)
plot(fitcv)
coefficient <- coef(fitcv, s="lambda.1se")

Active.index <- which(as.numeric(coefficient) != 0)
activa.coefficents <- as.numeric(coefficient)[Active.index]
siggenes <- rownames(coefficient)[Active.index]
# "B3GNT5"    "B3GNT9"    "C1GALT1C1" "CHST2"     "FUT9"      "GALNT13" 
print(siggenes)

```

### Using cox proportional hazards model to construct the equation of the marker genes

```{r}

formula_for_multi <- as.formula(paste0('Surv(OS,Status)~',paste(siggenes,sep='',collapse = '+')))
dt$Status = as.double(dt$Status)
multi_variate_cox <- coxph(formula_for_multi, data=dt)

# cox.zph Test the proportional hazards assumption for a Cox regression model fit 
ph_hypo_multi <- cox.zph(multi_variate_cox)
ph_hypo_table <- ph_hypo_multi$table[-nrow(ph_hypo_multi$table),]
formula_for_multi <- as.formula(paste0('Surv(OS,Status)~',paste(rownames(ph_hypo_table)[ph_hypo_table[,3]>0.05],sep='',collapse = '+')))
multi_variate_cox2 <- coxph(formula = formula_for_multi,data=dt)
candidate_genes <- c(rownames(ph_hypo_table)[ph_hypo_table[,3]>0.05])

# Forest
ggforest(model=multi_variate_cox2, data=dt,main = "Hazard ratios of Glioma candidates genes",fontsize=1)
# C_index <- multi_variate_cox2$concordance['concordance'] # accuracy
```


```{r}
# As for Glioma
riskscore <- function(survival_cancer_df, candidate_genes, cox_report){
  risk_score_table <- survival_cancer_df[,candidate_genes]
  for(gene_name in colnames(risk_score_table)){
    risk_score_table$gene <- risk_score_table[,gene_name]*(summary(cox_report)$coefficients[gene_name,1])
  }
  risk_score_table <- cbind(risk_score_table,"risk_score"=rowSums(risk_score_table)) %>%
    cbind(survival_cancer_df[,c("OS","Status")])
  risk_score_table <- risk_score_table[,c("OS","Status",candidate_genes,"risk_score")]
  }

# Create the final risk score tabel
risk_score_table <- riskscore(dt,candidate_genes, multi_variate_cox2)
```

# Time-dependent ROC to get the cutoff value

```{r}
TimeROC <- function(risk_score_table,times){
    time_roc_res <- timeROC(
      T = risk_score_table$OS,
      delta = risk_score_table$Status,
      marker = risk_score_table$risk_score,
      cause = 1,
      weighting="marginal",
      times = times,
      ROC = TRUE,
      iid = TRUE
    )
    
    time_ROC_df <- data.frame(
      TP_1 = time_roc_res$TP[, 1],
      FP_1 = time_roc_res$FP[, 1],
      TP_2 = time_roc_res$TP[, 2],
      FP_2 = time_roc_res$FP[, 2],
      TP_3 = time_roc_res$TP[, 3],
      FP_3 = time_roc_res$FP[, 3],
      TP_4 = time_roc_res$TP[, 4],
      FP_4 = time_roc_res$FP[, 4]
    )
    
    # TimeROC curves for 2, 5, 8 ,15 years were plotted in the model group. 
    # The best cutoff value was marked on the 5-year TimeROC curve.
    gg<-ggplot(data = time_ROC_df) +
      geom_line(aes(x = FP_1, y = TP_1), size = 0.7, color = "#3120E0") +
      geom_line(aes(x = FP_2, y = TP_2), size = 0.7, color = "#256D85") +
      geom_line(aes(x = FP_3, y = TP_3), size = 0.7, color = "#3B9AE1") +
      geom_line(aes(x = FP_4, y = TP_4), size = 0.7, color = "#21E1E1") +
      geom_abline(slope = 1, intercept = 0, color = "grey", size = 1, linetype = 2) +
      theme_bw() +
      annotate("text", x = 0.87, y = 0.23, size = 3.5,
               label = paste0(paste("AUC at "," years = ", sep = as.character(times[1])), 
                              sprintf("%.3f", time_roc_res$AUC[[1]])), color = "#3120E0"
      ) +
      annotate("text", x = 0.87, y = 0.16, size = 3.5,
               label = paste0(paste("AUC at "," years = ", sep = as.character(times[2])), 
                              sprintf("%.3f", time_roc_res$AUC[[2]])), color =  "#256D85"
      ) +
      annotate("text", x = 0.87, y = 0.09, size = 3.5,
               label = paste0(paste("AUC at "," years = ", sep = as.character(times[3])),  
                              sprintf("%.3f", time_roc_res$AUC[[3]])), color ="#3B9AE1"
      ) + 
      annotate("text", x = 0.875, y = 0.02, size = 3.5,
               label = paste0(paste("AUC at "," years = ", sep = as.character(times[4])), 
                              sprintf("%.3f", time_roc_res$AUC[[4]])), color =  "#21E1E1"
      ) +
      labs(x = "1-Specificity", y = "Sensitivity") +
      theme(
        # axis.text = element_text(face = "bold", size = 11, color = "black"),
        axis.title.x = element_text( size = 14, color = "black", margin = margin(c(15, 0, 0, 0))),
        axis.title.y = element_text( size = 14, color = "black", margin = margin(c(0, 15, 0, 0)))
      )
    
     cutoff <- risk_score_table$risk_score[which.max(time_ROC_df$TP_2-time_ROC_df$FP_2)]
     out <- list(gg,cutoff)
     
     return(out)
}


out<-TimeROC(risk_score_table,times = c(2, 5, 8,15))
AUC1 <- out[1]
cutoff <- out[2]
```

### Testing the model on the TCGA training dataset and testing dataset

```{r}
# As for training data
group <- (risk_score_table$risk_score > cutoff)
group[group == TRUE] <- 'high'
group[group == FALSE] <- 'low'
result_risk_table <- cbind(risk_score_table,group)

fit_km1 <- survfit(Surv(OS,Status) ~ group, data = 
                    result_risk_table)

# plot the result
gg1 <- ggsurvplot(fit_km1,risk.table = T,conf.int = F,legend.title="total risk score",pval = T,risk.table.height=0.3,xlab="year",title=paste0("TCGA Training Data"))
gg1$table <- gg1$table + theme(plot.title = element_text(size = 6, face = "bold"))
```

```{r}
# As for testing data
testing_score_table <- riskscore(pptesting,candidate_genes,multi_variate_cox2 )

group <- (testing_score_table$risk_score > cutoff)
group[group == TRUE] <- 'high'
group[group == FALSE] <- 'low'

result_risk_table <- cbind(testing_score_table,group)

fit_km2 <- survfit(Surv(OS,Status) ~ group, data = 
                    result_risk_table)

# plot the result
gg2 <- ggsurvplot(fit_km2,risk.table = T,conf.int = F,legend.title="total risk score",pval = T,risk.table.height=0.3,xlab="year",title=paste0("TCGA Testing Data"))
gg2$table <- gg2$table + theme(plot.title = element_text(size = 6, face = "bold"))
gg2
```

### Test the riskscore model on the CGGA dataset

```{r}
### Combine the 325 and 693 genes together
# CGGA325 <- file.path(data_dir,"CGGA.mRNAseq_325.txt")
# CGGA693 <- file.path(data_dir,"CGGA.mRNAseq_693.txt")
# cgga1 <- read.table(CGGA325,row.names = 1,header = T)
# cgga2 <- read.table(CGGA693,row.names = 1,header = T)
# 
# #change Fpkm to TPM
# FPKM2TPM <- function(fpkm){
#   exp(log(fpkm) - log(sum(fpkm)) + log(1e6))
# }
#
# TPMs1 <- apply(cgga1,2,FPKM2TPM)
# cgga1 <- as.data.frame(TPMs1) 
# TPMs2 <- apply(cgga2,2,FPKM2TPM)
# cgga2 <- as.data.frame(TPMs2) 
# 
# cgga1 <- cgga1[(rownames(cgga1) %in% candidate_genes),]
# cgga2 <- cgga2[(rownames(cgga2) %in% candidate_genes),]
# cgga1 <- as.data.frame(t(cgga1))
# cgga2 <- as.data.frame(t(cgga2))
# result <- rbind(cgga1,cgga2)
# save_path <- file.path(data_dir,"CGGA_candidate_genes.csv")
# write.table(result, file = save_path, sep = ",",quote=FALSE,row.names = T)

CGGA <- file.path(data_dir,"CGGA_candidate_genes.csv")
cgga_risk_table <- read.table(CGGA,row.names = 1,header = T,sep=",")
pp = preProcess(cgga_risk_table, method=c("scale"))
cgga_risk_table = predict(pp,cgga_risk_table)

# # B3GNT5*0.1906201+B3GNT9*0.3708307 + C1GALT1C1*0.2141980+ FUT9 *(-0.7967398)
cgga_risk_table$risk_score <- 0.1906201*cgga_risk_table$B3GNT5 + 0.3708307*cgga_risk_table$B3GNT9 + 0.2141980*cgga_risk_table$C1GALT1C1+(-0.7967398)*cgga_risk_table$FUT9

# combine the 325 and 693 clinical data together
# clinical1 <- file.path(data_dir,"CGGA.mRNAseq_325_clinical.txt")
# clinical2 <- file.path(data_dir,"CGGA.mRNAseq_693_clinical.txt")
# cgga_clinical1 <- read.table(clinical1,sep = "\t",row.names = 1,header = T)
# cgga_clinical2 <- read.table(clinical2,sep = "\t",row.names = 1,header = T)
# cgga_clinical <- rbind(cgga_clinical1,cgga_clinical2)
# clinical_all <- file.path(data_dir, "CGGA_clinical_all.csv")
# write.table(cgga_clinical, file =clinical_all, sep = ",",quote=FALSE,row.names = F)

clinical <- file.path(data_dir,"CGGA_clinical_all.csv")
cgga_clinical <- read.table(clinical,header = T,sep=",")
# cgga_clinical<-na.omit(cgga_clinical,c(cgga_clinical$OS,cgga_clinical$Censor))

cgga_risk_table$OS <- cgga_clinical$OS/356
cgga_risk_table$Status <- cgga_clinical$Censor
cgga_risk_table$type <-cgga_clinical$PRS.type
cgga_risk_table <- na.omit(cgga_risk_table)

# Plot the time dependent ROC of the CGGA
out2 <- TimeROC(cgga_risk_table,times = c(2, 5, 7, 10))
AUC2 <- out2[1]
cgga_cutoff <-out2[2]

group <- (cgga_risk_table$risk_score > cgga_cutoff)
group[group == TRUE] <- 'high'
group[group == FALSE] <- 'low'

cgga_result_risk_table <- cbind(cgga_risk_table,group)
fit_km3 <- survfit(Surv(OS,Status) ~ group, data = 
                    cgga_result_risk_table )

# plot the result
gg3 <- ggsurvplot(fit_km3, risk.table = T,conf.int = F,legend.title="total risk score",pval = T,risk.table.height=0.3,xlab="year")
gg3$table <- gg3$table + theme(plot.title = element_text(size = 6, face = "bold"))
```
### Show the result of CGGA according to the primary and recurrent glioma

```{r}
### As for the primary Glioma
primary_risk_table <- cgga_risk_table[cgga_risk_table$type=="Primary",]

group <- (primary_risk_table$risk_score > cgga_cutoff)
group[group == TRUE] <- 'high'
group[group == FALSE] <- 'low'

primary_risk_table <- cbind(primary_risk_table,group)
fit_km4 <- survfit(Surv(OS,Status) ~ group, data = 
                    primary_risk_table )

# plot the result
gg4 <- ggsurvplot(fit_km4, risk.table = T,conf.int = F,legend.title="total risk score",pval = T,risk.table.height=0.3,xlab="year",title=paste0("Primary Glioma"))
gg4$table <- gg4$table + theme(plot.title = element_text(hjust = 0.1, size = 6, face = "bold"))


### As for the recurrent Glioma
recurrent_risk_table <- cgga_risk_table[cgga_risk_table$type=="Recurrent",]
group <- (recurrent_risk_table$risk_score > cgga_cutoff)
group[group == TRUE] <- 'high'
group[group == FALSE] <- 'low'

recurrent_risk_table <- cbind(recurrent_risk_table,group)
fit_km4 <- survfit(Surv(OS,Status) ~ group, data = 
                    recurrent_risk_table )

# plot the result
gg4 <- ggsurvplot(fit_km4, risk.table = T,conf.int = F,legend.title="total risk score",pval = T,risk.table.height=0.3,xlab="year",title=paste0("Recurrent Glioma"))
gg4$table <- gg4$table + theme(plot.title = element_text(hjust = 0.1, size = 6, face = "bold"))
```

